Social Network Analysis: Week 3: Degree and Centrality

### 2022-02-08 docs/2. SNA-Tutorials/week2-SNA_Tutorial.Rmd

---
title: "SNA Tutorials: Social Overview"
output: html_notebook
---

@@

## Key Terms & Definitions
### Network Structure: Week 2: Key Terms
A *vertex attribute* is a vector (or variable) of additional information about the nodes in a network dataset:
name, location, demographic information about the actors or organizations  comprise the vertices of the network.

@@

*edge attribute* is a vector (or variable) of additional information about the edges or relationships in a network dataset. The most common edge attribute is weight, in a weighted dataset.

# Week 1: *Describe Network Data*
## 1. **List and inspect:** 
    List the objects to make sure the datafiles are working properly:
```is()
```

## 2. **Network Size:**
    What is the size of the network?

@@

## 3. **Network Features:**

@@

## 4. **Network Attributes:**
  R commands used with vector objects also work with network attribute data
```sort()
```
```summary()
```

@@

# Week2: *Network Structure: Overview*
- Dyads and Dyad Census
  *dyad census* gives the number of dyads where the relationship is:
    - Reciprocal (mutual), or mut
    - Asymmetric (non-mutual), or asym, and
    - Absent, or null

## Dyad Census 
  Classifies all dyads in the network as:
      - Reciprocal (mutual), or mut
      - Asymmetric (non-mutual), or asym, and
      - Absent, or null

@@

## **Triad Census**
  Classifies all triads in the network

@@

## Transitivity: **Global Clustering**

@@

**[statnet]**
  weighted network transitivity:
``` sna::gtrans()
```
  For Directed Networks - network structure rooted in hierarchy
```gtrans()
```

## Transivity: **Local Clustering**

@@

# *INCOMPLETE* Advanced [Week 2]
## ADVANCED: Path Length and Geodesic
## ADVANCED Component Structure and Membership

>>>>

---
title: "SNA Tutorials: Week 2: Network Structure"
output: html_notebook
---

@@

## Key Terms & Definitions
### Vetex Attribute & Edge Attribute 
A *vertex attribute* is a vector (or variable) of additional information about the nodes in a network dataset:
name, location, demographic information about the actors or organizations  comprise the vertices of the network.

@@

*edge attribute* is a vector (or variable) of additional information about the edges or relationships in a network dataset. The most common edge attribute is weight, in a weighted dataset.
### Network Structure
- Dyads and Dyad Census
  *dyad census* gives the number of dyads where the relationship is:
    - Reciprocal (mutual), or mut
    - Asymmetric (non-mutual), or asym, and
    - Absent, or null

# Network Structure [Week 2]
aspects of network structure introduced in the Week 2 Lecture:
- Dyads and Dyad Census
  *dyad census* gives the number of dyads where the relationship is:
    - Reciprocal (mutual), or mut
    - Asymmetric (non-mutual), or asym, and
    - Absent, or null

# Describe Network Data 
## 1. *List and inspect:* 
    List the objects to make sure the datafiles are working properly:
```is()
```

## 2. *Network Size:*
    What is the size of the network?

@@

## 3. *Network Features:*

@@

## 4. *Network Attributes*
  R commands used with vector objects also work with network attribute data
```sort()
```
```summary()
```

@@

# Dyad Census 
  Classifies all dyads in the network as:
      - Reciprocal (mutual), or mut
      - Asymmetric (non-mutual), or asym, and
      - Absent, or null

@@

# Triad Census
  Classifies all triads in the network

@@

# Transitivity or Global Clustering

@@

**[statnet]**
  weighted network transitivity:
``` sna::gtrans()
```
  For Directed Networks - network structure rooted in hierarchy
```gtrans()
```
# Local Transivity or Clustering

@@

# Advanced [Week 2]
## ADVANCED: Path Length and Geodesic
## ADVANCED Component Structure and Membership

### 2022-02-08 docs/3. SNA - Assignments/Week3-Assignment.Rmd

```

      *Week 3 Assignment: Degree and Centrality*

Using the tutorial as a guide, 
      *  you may use your own data; 
      *  the data provided in the week 3. Syntax dataset is Correlates of War trade network - you can read more about it online; or
      *  choose another dataset.

## Working from the example R syntax:
1.   calculate network density; 
2.   degree centrality, and 
3.   centralization measures for your network data.

@@

##
1.  [**Identify an existing data set.**]{.ul}

| This can be one provided in the course directory, in an R package or library, located online, or some other source.

2.  [**Briefly describe the network dataset.**]{.ul}

| Identify and describe content of nodes and links, and identify format of data set (i.e., matrix or edgelist, directed or not, weighted or not), and whether attribute data are present. 
  Be sure to provide information about network size (e.g., information obtained from network description using week 1 network basics tutorial commands.)

# Week 2: Scripts

```{r Import Data Set, echo=TRUE}

  # install.packages("network")

    data("flo", package = "network")

```

# 2. Briefly describe the network dataset.

-   *Be sure to provide information about network size (e.g., information obtained from network description using week 1 network basics tutorial commands.)*

1.  *Identify and describe content of nodes and links,*

-   each of the 16 Florentine Family Names represent a Node in the network.

```{r Colnames}
 colnames(flo)
```

2.  *Identify format of data set (i.e., matrix or edgelist, directed or not, weighted or not), and*

    -   Flo Format: Size: 16 x 16 square dimmentions indicate a adjacency matrix format

    Network attributes: vertices = 16 directed = TRUE hyper = FALSE loops = FALSE multiple = FALSE bipartite = FALSE total edges= 40 missing edges= 0 non-missing edges= 40

Vertex attribute names: vertex.names

No edge attributes

```{r Dimmentions, echo=TRUE}

  dim(flo)

```

```{r Network Attributes, echo=TRUE}

  network(flo)

```

3.  *Identify whether attribute data are present.*

```{r Summary,echo=TRUE}

  summary(flo,
          print.adj = TRUE
          )
```

## Create & Describe: Directed Adjacency Matrix

*Florentine Families Directed Adjacency Matrix Network Objects:*
    **iGraph** flo.ig 
    **Statnet** flo.stat

iGraph and Statnet both use {r graph4} for Directed Adjacency Matrix Objects

```{r graph4, statnet: Directed: Adjacency Network, echo=TRUE}

    flo.stat<-
        network(
        flo,
        directed=F,
        matrix.type="adjacency"
        )

#     provides a description of several critical network features
    print(flo.stat)

```

```{r graph4, igraph: Directed: Adjacency Network, echo=TRUE}

    flo.ig<-
      graph.adjacency(
      flo,
      mode="undirected"
      ) 
    print(flo.ig)
```

```{r igraph: Network Description, echo=TRUE}

#     Count Vertices
    vcount(flo.ig)

#     Count Edges
    ecount(flo.ig)

# Is this a Bipartite or single mode network?
    is_bipartite(flo.ig)

#  Are edges directed or undirected?
    is_directed(flo.ig)

#Are edges weighted or unweighted?
    is_weighted(flo.ig)

```
 **Vertex and Edge Attributes**

```{r statnet: Attribute Data: Directed: Adjacency Network, echo=TRUE}

#     access vertex attributes
  head(flo.stat %v% "vertex.names")

#      list the names of vertex attributes
  network::list.vertex.attributes(flo.stat)

#     access edge attribute
  head(flo.stat%e% "weight")

#  Generate a list the names of edge attributes
  network::list.edge.attributes(flo.stat)

#   summarize numeric network attribute
  summary(flo.stat  %v% "name")

```

```{r igraph: Attribute Data: Directed: Adjacency Network, echo=TRUE}

#     access vertex attribute 
  V(flo.ig)$name
#     Generate a list the names of vertex attributes
  igraph::vertex_attr_names(flo.ig)

#     access edge attribute
  E(flo.ig)$weight

#   Generate a list the names of edge attributes  
  igraph::edge_attr_names(flo.ig)

#   summarize numeric network attribute
  summary(E(flo.ig)$weight)

```

# 3. Explore the dataset using commands from week 2 tutorial.

5.  *geodesic and path distances,*

## Path Length and Geodesic

```{r iGraph:Nodes: Path: Length, echo=TRUE}

#     Calculate distances between two nodes
    distances(flo.ig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(flo.ig,"Bischeri", "Castellani",weights=NA)

```

We can also find all of the shortest paths between two famalies in this network.
These list the starting node and the ending node with all nodes inbetween for each path. 
The option weights=NA means that any available edge weights are ignored.

```{r iGraph:Nodes: Path: Shortest, echo=TRUE}

#     isolating the distances between specific node
    all_shortest_paths(flo.ig,"Strozzi","Tornabuoni", weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

The concept of shortest path to describe the overall network structure can be more useful

```{r iGraph:Network: Path: Shortest, echo=TRUE}

#     find average shortest path for network
    average.path.length(flo.ig,directed=F)

```

6.  *triads or transitivity,*

## Dyad Census

Classifies all dyads in the network as: - Reciprocal (mutual), or mut - Asymmetric (non-mutual), or asym, and - Absent, or null

```{r statnet: Dyad Census , echo=TRUE}

    sna::dyad.census(flo.stat)

```

```{r iGraph: Dyad Census, echo=TRUE}

  igraph::dyad.census(flo.ig)

```

## Triad Census

Classifies all triads in the network. The triad census provides a fundamental descriptive insight into the types of triads found in a particular dataset.

```{r statnet: Triad Census, echo=TRUE}

#     Classify all Triad in the network: Directed Network
  sna::triad.census(flo.stat, mode= )

#     Classify all Triad in the network: Undirected Network
   sna::triad.census(flo.stat, mode="graph")

#     total number of all four triad types returned by triad census
   sum(sna::triad.census(flo.stat, mode="graph"))

```

```{r iGraph: Triad Census, echo=TRUE}
#Classify all Triad in the network
     igraph::triad_census(flo.ig)

```

## Transitivity or Global Clustering

```{r iGraph: weighted/ network transitivity, echo=TRUE}

#     network transitivity:
    transitivity(flo.ig)

#     weighted network transitivity:
#   igraph::transivity(flo.ig)

```

```{r statnet: weighted/ directed network transitivity, echo=TRUE}

#     Directed Networks transitivity:
  gtrans(flo.stat)

#   methods for weighted networks, rank and correlation, along with the relevant references.
#  sna::gtrans()

```

## Local Transivity or Clustering
`Local transitivity` (local clustering coefficient), is a technical description of the density of an ego network.

```{r iGraph:Local transitivity, echo=TRUE}

#     Retrive a list of the vertices we are interested
    V(flo.ig)[c("Bischeri" ,
                "Castellani", 
                "Medici",
                "Ridolfi")]

#     check ego network transitivity
transitivity(flo.ig,
             type="local", 
             vids=V(flo.ig)
             [c("Bischeri",
                "Castellani", 
                "Medici",
                "Ridolfi")]) 
#     get global clustering cofficient
  transitivity(flo.ig, type="global")

#     get average local clustering coefficient
  transitivity(flo.ig, type="average")

```

7.  *connectedness and.or component structure, etc.*

> Distill is a publication format for scientific and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.
##

# Week3:    Briefly describe the dataset:
1.   identify initial network format; 
2.   describe and identify the nodes 
     including: 
        *  how many nodes are in the dataset; 
3.   what constitutes a tie or edge
     including:
        *  how many ties, 
        *   whether ties are directed/undirected,
        *   whether ties are weighted/binary, 
        *   and how to interpret the value of the tie if any.
        *   Whether or not there are edge attributes 
            that might be used to:
            *   subset data, or 
            *   stack multiple networks.

#WEEK 3:     Working from the example R syntax:
1.   calculate network density; 
2.   degree centrality, and 
3.   centralization measures for your network data.

#WEEK 3:  Provide at least two or three noteworthy results
1.   the relevant statistics and interpretation. 
        For example, explaining which node(s) are most central and which are least central. 
2.   Discuss (with any related evidence) whether or not the node(s) behavior is in line with or violates expectations based on the degree centrality measure. 
3.   What do you make of network density and centralization measures?  
4.   What about differences in the binary 100 million threshold network vs valued trade flows network. 
 5.  Check tutorial for other ideas of results to discuss.

>>>>

```
## Working from the example R syntax:
1.   calculate network density; 
2.   degree centrality, and 
3.   centralization measures for your network data.

@@

> Distill is a publication format for scientific and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.
